Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2017-01-25 manuscript/reviewer_response_tree_physiology.Rmd

We are grateful for the opportunity to resubmit our manuscript. We thank the reviewers for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from both referees.

### **Referee #1**

\
- *Reviewer: 1. The role of nutrient defiency may have been underestimated...The relative contribution should be determined...*

\
*Response*: We were successfully able to identify not one, but two processes that likely contributed to photosynthetic inhibition. We agree that determining the contribution of both of the processes, nutrient limitation and starch accumulation, is important as they are addressed at length in the discussion. The full model including both nitrogen and starch explains 36% of the variation in *A*~max~ on a mass basis. Including only nitrogen in the model explains 25% and with only starch explains 11% of the variation. The interaction between nitrogen and starch explained practically no variation. Consequently, nutrient deficiency composes 69% of the explained variation and starch accumulation the rest. We have added these new results in the "Gas exchange and photosynthetic parameters" section of the Results. By providing convincing empirical evidence of these processes, this experimental design should provide a framework for future work to further quantify the contributions of each.

### 2017-01-25 manuscript/manuscript_pve.Rmd

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

##Do reductions in photosynthesis explain reductions in seedling growth?
We use a simple approach to estimate cumulative seedling C gain with measured reductions in leaf *A*~n~ , via soil volume effects.  Although reductions in both *A*~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter *et al.* [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. As estimated whole-plant C use efficiency was decreased in seedlings with belowground sink limitation, this suggests that other components of the C balance were affected. These results are especially noteworthy for process-based growth models that parameterize CUE, tissue respiration and C allocation as fixed processes. For example, classical approaches in tree growth and production modelling are often driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible these results represent an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *V*~cmax~, *R*~dark~ and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in C-balance driven plant growth models [@lacointe2000carbon].

>>>>

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf N~f~ and leaf starch with *A*~max~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). Leaf N~f~ and leaf starch explained 25% and 11% of the variation in *A*~max~ on a mass basis, respectively. The interaction between leaf N~f~ and leaf starch explained practically no variation in *A*~max~ on a mass basis. We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

##Do reductions in photosynthesis explain reductions in seedling growth?
We use a simple approach to estimate cumulative seedling C gain with measured reductions in leaf *A*~n~, via soil volume effects.  Although reductions in both *A*~max~ and biomass were strongly correlated among treatments, as hypothesized by Poorter *et al.* [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. As estimated whole-plant C use efficiency was decreased in seedlings with belowground sink limitation, this suggests that other components of the C balance were affected. These results are especially noteworthy for process-based growth models that parameterize CUE, tissue respiration and C allocation as fixed processes. For example, classical approaches in tree growth and production modelling are often driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible these results represent an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *V*~cmax~, *R*~dark~ and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in C-balance driven plant growth models [@lacointe2000carbon].
